---
title: "CAP study: Add scored qualtrics data to AX and RDM clean datasets"
author: "Hayley Brooks"
---

Output:
1) AXallClean_Qualtrics.csv
2) RDMallClean_Qualtrics.csv


```{r}
# clean environment
rm(list=ls());

# configuration
config = config::get()

# load data
scoredQualtricsPath = file.path(config$path$combined, config$QUALTRICScsvs$Combined_subID_scored_noDuplicates)
scoredQualtrics = read.csv(scoredQualtricsPath)
#scoredQualtrics = read.csv("/Volumes/CAP/data/combinedData/QualtricsCombined_subID_scored_noDuplicates.csv"); # scored qualtrics

rdmDFpath = file.path(config$path$combined, config$RDMcsvs$RDM_allClean);
rdmDF = read.csv(rdmDFpath);
#rdmDF = read.csv("/Volumes/CAP/data/combinedData/RDMallClean.csv"); # RDM data

axDFpath = file.path(config$path$combined, config$AXcsvs$AX_allClean);
axDF = read.csv(axDFpath);
#axDF = read.csv("/Volumes/CAP/data/combinedData/AXallClean.csv"); # AX data   

subNum = unique(rdmDF$subID);
nSub = length(unique(rdmDF$subID)); #544

```


We can pair down scoredQualtrics to make it easier to work with, lets keep the columns we want

1) stai_s_score (column 33)
2) stai_t_score (column 56)
3) pss_score (column 67)
4) pss_stressedToday (column 70)
5) uclal_score (column 91)
6) columns 92:140 (e.g. covid qs, location, etc)
```{r}
qualtricsDF = scoredQualtrics[,c(33,56,67,70,91:140)]; # keep only the columns we listed above
```

Merge the data frames by sub ID and phase (this will also get rid of the redundant columns subID and phase):
```{r}
AXallClean_Qualtrics = merge(axDF,qualtricsDF,by=c("subID","phase"), all.x = T); # AX
RDMallClean_Qualtrics = merge(rdmDF,qualtricsDF,by=c("subID","phase"), all.x = T, all.y = T);  # RDM
```

Save the output:
```{r}

AXoutputpath = file.path(config$path$combined, config$AXcsvs$AX_qualtrics);
RDMoutputpath = file.path(config$path$combined, config$RDMcsvs$RDM_qualtrics);

write.csv(file=AXoutputpath, AXallClean_Qualtrics);
write.csv(file=RDMoutputpath, RDMallClean_Qualtrics);
#write.csv(file = "/Volumes/CAP/data/combinedData/AXallClean_Qualtrics.csv", AXallClean_Qualtrics);
#write.csv(file = "/Volumes/CAP/data/combinedData/RDMallClean_Qualtrics.csv", RDMallClean_Qualtrics);

```


When you save the notebook, an HTML file containing the code and output will be saved alongside it (click the *Preview* button or press *Cmd+Shift+K* to preview the HTML file). 

The preview shows you a rendered HTML copy of the contents of the editor. Consequently, unlike *Knit*, *Preview* does not run any R code chunks. Instead, the output of the chunk when it was last run in the editor is displayed.

