---
title: "CAP study: Add scored qualtrics and quartiles data to AX and RDM clean datasets"
author: "Hayley Brooks"
---

Output:
1) AXallClean_Qualtrics.csv
2) RDMallCleanGain_Qualtrics.csv
3) RDMallCleanLoss_Qualtrics.csv

```{r}
# clean environment
rm(list=ls());

# configuration
config = config::get()

# load data
scoredQualtricsPath = file.path(config$path$combined, config$QUALTRICScsvs$Combined_subID_scored_noDuplicates); # scored qualtrics
scoredQualtrics = read.csv(scoredQualtricsPath);

stateQuartile_csv = file.path("../documentation",config$documentationFiles); # not sure if this will work for everyone but if you're located in the GitHub 'cap/preprocessing/' directory, then this will take you to the cap/documentation directory where the stateQuartile.csv file is located.




rdmDFpath = file.path(config$path$combined, config$RDMcsvs$RDM_allClean); # RDM dataset (both gain and loss tasks)
rdmDF = read.csv(rdmDFpath);
stateQuartile = read.csv(stateQuartile_csv); # load state and quartile information


axDFpath = file.path(config$path$combined, config$AXcsvs$AX_allClean); # AX dataset
axDF = read.csv(axDFpath);
 

subNum = unique(rdmDF$subID);
nSub = length(unique(rdmDF$subID)); #544

```

```{r}
library(plyr) # need this for 'join' function
```


We can pair down scoredQualtrics to make it easier to work with, lets keep the columns we want

1) stai_s_score (column 120)
2) stai_t_score (column 119)
3) pss_score (column 121)
4) pss_stressedToday (column 65)
5) uclal_score (column 122)
6) columns 86:114 (e.g. covid qs, location, etc)
7) columns 117:141 (^ continued)
```{r}
qualtricsDF = scoredQualtrics[,c(65,86:114,117:141)]; # keep only the columns we listed above
```

Merge the data frames by sub ID and phase (this will also get rid of the redundant columns subID and phase):
```{r}
#AXallClean_Qualtrics = base::merge(axDF,qualtricsDF, by=c("subID","phase"), all.x = T); # AX
#RDMallClean_Qualtrics = base::merge(rdmDF,qualtricsDF, by=c("subID","phase"), all.x = T, all.y = T);  # RDM

AXallClean_Qualtrics = join(axDF, qualtricsDF, by=c("subID", "phase"))
RDMallClean_Qualtrics = join(rdmDF,qualtricsDF, by=c("subID","phase"));  # RDM

# using the 'merge' function merges the datasets but rearranges the rows within each participant which is not great if we want to do to temporal stuff. using the join function allows us to merge the datasets and preserve row order.

```



Add quartile information based on "loc_state"
```{r}

#RDM
tmpQuartileVec = RDMallClean_Qualtrics$loc_state; # create temporary vector and store state info
tmpQuartileVecNumeric = vector(length = length(tmpQuartileVec)); # numerical vector to store values 1-4 for quartiles since tmpQuartileVec is table with factors

tmpQuartileVecNumeric[tmpQuartileVec %in% stateQuartile$Quartile.1] = 1; # quartile 1
tmpQuartileVecNumeric[tmpQuartileVec %in% stateQuartile$Quartile.2] = 2; # quartile 2
tmpQuartileVecNumeric[tmpQuartileVec %in% stateQuartile$Quartile.3] = 3; # quartile 3
tmpQuartileVecNumeric[tmpQuartileVec %in% stateQuartile$Quartile.4] = 4; # quartile 4
tmpQuartileVecNumeric[tmpQuartileVec == "District of Columbia"] = 2; # DC is "Columbia" in stateQuartile csv

RDMallClean_Qualtrics$quartile = tmpQuartileVecNumeric; # everyone should have 1-4 for quartile except for sub ID 461 phase 1 bc we don't have qualtrics responses for this person.

#AX
tmpQuartileVec = AXallClean_Qualtrics$loc_state; # create temporary vector and store state info
tmpQuartileVecNumeric = vector(length = length(tmpQuartileVec)); # numerical vector to store values 1-4 for quartiles since tmpQuartileVec is table with factors

tmpQuartileVecNumeric[tmpQuartileVec %in% stateQuartile$Quartile.1] = 1; # quartile 1
tmpQuartileVecNumeric[tmpQuartileVec %in% stateQuartile$Quartile.2] = 2; # quartile 2
tmpQuartileVecNumeric[tmpQuartileVec %in% stateQuartile$Quartile.3] = 3; # quartile 3
tmpQuartileVecNumeric[tmpQuartileVec %in% stateQuartile$Quartile.4] = 4; # quartile 4
tmpQuartileVecNumeric[tmpQuartileVec == "District of Columbia"] = 2; # DC is "Columbia" in stateQuartile csv

AXallClean_Qualtrics$quartile = as.numeric(tmpQuartileVecNumeric); # everyone should have 1-4 for quartile except for sub ID 461 phase 1 bc we don't have qualtrics responses for this person.

```


Age was only collected in phase 1, lets fill in age for phase 2 demo_age variable based on phase 1 response
```{r}

#AX
for(s in 1:nSub){
  subData = AXallClean_Qualtrics[AXallClean_Qualtrics$subID==subNum[s],]
  
  if (length(unique(subData$phase))>1){ #if we have phase 2 data for this participant
    
    AXallClean_Qualtrics$demo_age[AXallClean_Qualtrics$subID==subNum[s] & AXallClean_Qualtrics$phase==2] = subData$demo_age[1]; # replace phase 2 demo_age with sub's age
  }
}

#RDM
for(s in 1:nSub){
  subData = RDMallClean_Qualtrics[RDMallClean_Qualtrics$subID==subNum[s],]
  
  if (length(unique(subData$phase))>1){ #if we have phase 2 data for this participant
    
    RDMallClean_Qualtrics$demo_age[RDMallClean_Qualtrics$subID==subNum[s] & RDMallClean_Qualtrics$phase==2] = subData$demo_age[1]; # replace phase 2 demo_age with sub's age
  }
}

```


Create separate datasets for the gain only and loss only tasks (but keep the original one as is: "RDMallClean_Qualtrics")
```{r}

RDMallCleanGain_Qualtrics = RDMallClean_Qualtrics[RDMallClean_Qualtrics$rdmTask==1,]; #107219 rows in gain dataset
RDMallCleanLoss_Qualtrics = RDMallClean_Qualtrics[RDMallClean_Qualtrics$rdmTask==2,]; #17353 rows in loss dataset

```



Save the output:
```{r}

AXoutputpath = file.path(config$path$combined, config$AXcsvs$AX_qualtrics);
RDMbothOutputpath = file.path(config$path$combined, config$RDMcsvs$RDMbothTasks_qualtrics);
RDMlossOutputpath = file.path(config$path$combined, config$RDMcsvs$RDMloss_qualtrics);
RDMgainOutputpath = file.path(config$path$combined, config$RDMcsvs$RDMgain_qualtrics);

write.csv(file=AXoutputpath, AXallClean_Qualtrics);
write.csv(file=RDMbothOutputpath, RDMallClean_Qualtrics);
write.csv(file=RDMlossOutputpath, RDMallCleanLoss_Qualtrics);
write.csv(file=RDMgainOutputpath, RDMallCleanGain_Qualtrics);


```



