---
title: "CAP study: Add scored qualtrics data to AX and RDM clean datasets"
author: "Hayley Brooks"
---

Output:
1) AXallClean_Qualtrics.csv
2) RDMallClean_Qualtrics.csv


```{r}
# clean environment
rm(list=ls());

# configuration
config = config::get()

# load data
scoredQualtricsPath = file.path(config$path$combined, config$QUALTRICScsvs$Combined_subID_scored_noDuplicates)
scoredQualtrics = read.csv(scoredQualtricsPath)
#scoredQualtrics = read.csv("/Volumes/CAP/data/combinedData/QualtricsCombined_subID_scored_noDuplicates.csv"); # scored qualtrics

rdmDFpath = file.path(config$path$combined, config$RDMcsvs$RDM_allClean);
rdmDF = read.csv(rdmDFpath);
#rdmDF = read.csv("/Volumes/CAP/data/combinedData/RDMallClean.csv"); # RDM data

axDFpath = file.path(config$path$combined, config$AXcsvs$AX_allClean);
axDF = read.csv(axDFpath);
#axDF = read.csv("/Volumes/CAP/data/combinedData/AXallClean.csv"); # AX data   

subNum = unique(rdmDF$subID);
nSub = length(unique(rdmDF$subID)); #544

```

```{r}
library(plyr) # need this for 'join' function
```


We can pair down scoredQualtrics to make it easier to work with, lets keep the columns we want

1) stai_s_score (column 33)
2) stai_t_score (column 56)
3) pss_score (column 67)
4) pss_stressedToday (column 70)
5) uclal_score (column 91)
6) columns 92:140 (e.g. covid qs, location, etc)
```{r}
qualtricsDF = scoredQualtrics[,c(33,56,67,70,91:140)]; # keep only the columns we listed above
```

Merge the data frames by sub ID and phase (this will also get rid of the redundant columns subID and phase):
```{r}
#AXallClean_Qualtrics = base::merge(axDF,qualtricsDF, by=c("subID","phase"), all.x = T); # AX
#RDMallClean_Qualtrics = base::merge(rdmDF,qualtricsDF, by=c("subID","phase"), all.x = T, all.y = T);  # RDM
AXallClean_Qualtrics = join(axDF, qualtricsDF, by=c("subID", "phase"))
RDMallClean_Qualtrics = join(rdmDF,qualtricsDF, by=c("subID","phase"));  # RDM

# using the 'merge' function merges the datasets but rearranges the rows within each participant which is not great if we want to do to temporal stuff. using the join function allows us to merge the datasets and preserve row order.

```

Save the output:
```{r}

AXoutputpath = file.path(config$path$combined, config$AXcsvs$AX_qualtrics);
RDMoutputpath = file.path(config$path$combined, config$RDMcsvs$RDM_qualtrics);

write.csv(file=AXoutputpath, AXallClean_Qualtrics);
write.csv(file=RDMoutputpath, RDMallClean_Qualtrics);
#write.csv(file = "/Volumes/CAP/data/combinedData/AXallClean_Qualtrics.csv", AXallClean_Qualtrics);
#write.csv(file = "/Volumes/CAP/data/combinedData/RDMallClean_Qualtrics.csv", RDMallClean_Qualtrics);

```



