---
title: 'CAP: RDM analysis'
author: "Hayley Brooks"
date: "5/13/2021"
output: html_document
---

Data set up is taken care of by RDMsteup.R script. 
```{r setup, include=FALSE}
rm(list=ls())

library('config')
config = config::get()


setup_source = file.path(config$code_files$setup_data)
source(setup_source) #, local = knitr::knit_global())


# Looks for existing model file and if its found, then load it and we will now have our previous models in the environment
outputfile = file.path(config$path$Rdata, config$Rdata_files$models)

#if(exists(outputfile)){
  load(outputfile)
#}

```

Plot an example choice set for the gain-only task
```{r}
plot(rdmGainQualtrics$rdmAlternative[rdmGainQualtrics$subID==1], col=rdmGainQualtrics$rdmGroundEV, pch=16, xlab="trial", ylab="safe ($)", main="Safe amounts across the task\ncolors = levels")

plot(rdmGainQualtrics$rdmRiskyGain[rdmGainQualtrics$subID==1], col=rdmGainQualtrics$rdmGroundEV, pch=16, xlab="trial", ylab="risky gain ($)", main="Risky gain amounts across the task\ncolors = levels")

plot(rdmGainQualtrics$rdmGroundEV[rdmGainQualtrics$subID==1], col=rdmGainQualtrics$rdmGroundEV, pch=16, xlab="trial", ylab="level or context ($)", main="Context created by shared expected values")


```
 


### Risky decision-making analysis
#### Probability of gambling GAIN ONLY TASK
```{r}
# GROUP LEVEL

Phs1Pgam = mean(rdmGainQualtrics$rdmChoice[rdmGainQualtrics$phase==1], na.rm=T); # phase 1 pgamble = 0.3901198 (n=518)
Phs2Pgam = mean(rdmGainQualtrics$rdmChoice[rdmGainQualtrics$phase==2], na.rm=T); # phase 2 pgamble = 0.3783025 (n=326)
bothPhsPgam = mean(rdmGainQualtrics$rdmChoice[rdmGainQualtrics$subID %in% BothPhsSubIDs], na.rm=T); # pgamble across participants who completed both phase 1 and phase 2 (n=313) = 0.3867135

# 95% CI
# phase 1 (n=518)
me1 = qnorm(.975)*(Phs1Pgam/sqrt(Phs1nSub)); #0.03277965
Phs1Pgam + me1; #0.4134254
Phs1Pgam - me1; #0.3478662

# phase 2 (n=326)
me2 = qnorm(.975)*(Phs2Pgam/sqrt(Phs2nSub)); #0.04032889
Phs2Pgam + me2; #0.4118445
Phs2Pgam - me2; #0.3311867

# phase 1 and phase 2 (n = 313)
me3 = qnorm(.975)*(bothPhsPgam/sqrt(BothPhsnSub)); #0.04192421
bothPhsPgam + me3; #0.4203572
bothPhsPgam - me3; #0.3365088


# INDIVIDUAL LEVEL

# data frame where each row is a participant and there are 6 columns: {subID, day, phase 1 pgam, phase 2 pgam, both phases pgam, diffPgam}. If a participant is not included because they did not complete phase 2 or they were excluded, there will be an NaN in the corresponding  column. "diffPgam" is the difference in probability of gambling between two phases if we have that information for a given participant

pGambleSubID = as.data.frame(matrix(data=NA, nrow =nSubB4exclusion, ncol=6, dimnames=list(c(NULL), c("subID", "day", "phs1pgam","phs2pgam","bothPhspgam", "diffPgam"))))

for (s in 1:nSubB4exclusion) {
  
  
  subPhs1 = rdmGainQualtrics[rdmGainQualtrics$subID==subNumB4exclusion[s] & rdmGainQualtrics$phase==1,rdmColumns];
  subPhs2 = rdmGainQualtrics[rdmGainQualtrics$subID==subNumB4exclusion[s] & rdmGainQualtrics$phase==2,rdmColumns];
  subDay = rdmGainQualtrics$day[rdmGainQualtrics$subID==subNumB4exclusion[s]]
  
  pGambleSubID$subID[s] = subNumB4exclusion[s]
  pGambleSubID$day[s] = subDay[1]
  pGambleSubID$phs1pgam[s] = mean(subPhs1$rdmChoice, na.rm=T)
  pGambleSubID$phs2pgam[s] = mean(subPhs2$rdmChoice, na.rm=T)
  pGambleSubID$bothPhspgam[s] = mean(c(subPhs1$rdmChoice, subPhs2$rdmChoice), na.rm=T)
  pGambleSubID$diffPgam[s] = pGambleSubID$phs2pgam[s] - pGambleSubID$phs1pgam[s]
 
};

# NA bothPhspgam for people who did not participate in phase 2!!

# plot histogram of the change in risk-taking between phases (diff > 0 = more risk taking phase 1; diff < 0 =more risk taking phase 2; diff = 0, similar risk taking across phases)
summary(pGambleSubID$diffPgam); # mean = -.0112 (indicating generally more risk taking in phase 2)
hist(pGambleSubID$diffPgam, main="Distribution of change in risk taking", xlab="change in risk-taking\n p(gamble phase 2) - p(gamble phase 1)", ylab="number of participants",breaks = 100);

par(pty = 's'); # make the figure a square
plot(pGambleSubID$phs1pgam, pGambleSubID$phs2pgam, abline(a=0, b=1), xlab="p(gamble phase 1)", ylab="p(gamble phase 2)", main="Risk-taking in both phases (mean in red)", asp=1, ylim=c(0,1), xlim=c(0,1));
points(mean(pGambleSubID$phs1pgam,na.rm=T), mean(pGambleSubID$phs2pgam,na.rm=T), col="red", pch=17);


```

#### Simple regressions for risky decision-making GAIN ONLY TASK
1) running on lab computer speeds it up a bit, still take a couple minutes.
2) Use glm to explore but very carefully (these take a second)
3) save the R data models (all model names start with "model" except those we don't want saved will start with "tmp")
 - each model has an if statement (using exist function) so that models we already have don't need to be run again

Note: adding "na.action = na.exclude" retains original row number which we need when pulling out residuals


### DOES TASK-RELATED STUFF ACCOUNT FOR RISK-TAKING (main model plus other variations we tried):
```{r}
# Simple, main model that we will take residuals from to do the remaining analyses (has task events only):
if(!exists('model_task')){
  model_task = glmer(rdmChoice~ 1 + rdmGainSC + rdmSafeSC + rdmGroundEV + (0 + rdmGainSC + rdmSafeSC|subID),data=rdmGainQualtrics,family="binomial",na.action=na.exclude, control=glmerControl(optimizer = "bobyqa",optCtrl = list(maxfun=1000000)) );
}
# AIC = 73534.3 - this is not the best AIC but we decided that it probably makes the most sense to keep this model simple (just letting task-related stuff do their things) and then see what else accounts for RDM behavior.

# Note: kif we put a 1 in the random effects, it could wipe out subject level effects that we are interested in like affective measures (it wouldn't get rid of trial by trial stuff tho)


# Other models that we tried for our first-level and aren't going to use for residual analysis but want to save:

# Task events with phase and day:
if(!exists('model_taskPhaseDay')){
  model_taskPhaseDay=glmer(rdmChoice~ 1 + rdmGainSC + rdmSafeSC + rdmGroundEV + phaseRecode + dayOverallSC + (0 + rdmGainSC + rdmSafeSC|subID),data=rdmGainQualtrics,family="binomial", na.action=na.exclude, control=glmerControl(optimizer = "bobyqa",optCtrl = list(maxfun=1000000)) );
}
# AIC = 73516.2
# less risk taking in phase 2
# no effect of day 


# Task events with just phase
if(!exists('model_taskPhase')){
  model_taskPhase=glmer(rdmChoice~ 1 + rdmGainSC + rdmSafeSC + rdmGroundEV + phaseRecode + (0 + rdmGainSC + rdmSafeSC|subID),data=rdmGainQualtrics,family="binomial",na.action=na.exclude, control=glmerControl(optimizer = "bobyqa",optCtrl = list(maxfun=1000000)) );
}
# AIC = 73516.4
# less risk taking in phase 2

```

### Store residuals from our main model (model_task):
```{r}
rdmGainQualtrics$predmTask = predict(model_task, type="link"); 
```

### Models that involve residuals (the bulk of the analysis):

```{r}
# STAIS, STAIT, UCLAL, PSS:
if(!exists('model_pred_stais')){
  model_pred_stais = glmer(rdmChoice~  0 + stai_s_score_scaled + (1|subID), data=rdmGainQualtrics, family="binomial", control=glmerControl(optimizer = "bobyqa",optCtrl = list(maxfun=1000000)), offset=predmTask );
}
# singular
# no effect of stais
# AIC = 68606.6
# nObs = 98951

if(!exists('model_pred_stait')){
  model_pred_stait = glmer(rdmChoice~  0 + stai_t_score_scaled + (1|subID), data=rdmGainQualtrics, family="binomial", control=glmerControl(optimizer = "bobyqa",optCtrl = list(maxfun=1000000)), offset=predmTask  );
}
# no effect of stait
# AIC = 68606.6
# nObs = 98951
# singular

if(!exists('model_pred_uclal')){
  model_pred_uclal = glmer(rdmChoice~  0 + uclal_score_scaled + (1|subID), data=rdmGainQualtrics, family="binomial", control=glmerControl(optimizer = "bobyqa",optCtrl = list(maxfun=1000000)) ,offset=predmTask  );
}
# no effect of uclal
# AIC = 68606.6
# nObs = 98951
# singular


if(!exists('model_pred_pss')){
  model_pred_pss = glmer(rdmChoice~ 0 + pss_score_scaled + (1|subID), data=rdmGainQualtrics, family="binomial", control=glmerControl(optimizer = "bobyqa",optCtrl = list(maxfun=1000000)),offset=predmTask )
}
# no effect of pss
# AIC = 68606.7
# nObs = 98951
# singular


# COVID RISK: 
if(!exists('model_pred_covidRisk')){
  model_pred_covidRisk= glmer(rdmChoice ~  0 + covq_PAB_q1_personalRisk_scaledNoNA  + (1|subID), data=rdmGainQualtrics, family="binomial", na.action=na.exclude, control=glmerControl(optimizer = "bobyqa",optCtrl = list(maxfun=1000000)), offset=predmTask ); 
}

# AIC = 68791.1 
# nObs = 99307
# effect of covid risk is negative and trending (beta = -.025(.015), p=.0998)
# it was very significant in the main models above (not using residuals) - possible that the effect will change when in a model with temporal context effects
# singular

# covid risk mean: 5.26 phase 1 and 5.04 in phase 2

# TO DO:
# covid risk variable is capturing across and within subject stuff:
# 1) resrict data to one phase and if the pattern holds up, we atleast know that individual differences do matter
# 2) relate changes in risk taking to changes in perceived risk (using a variable that captures the difference in covid risk across phase 1)

# Affective measures... 
# - No observed effects (at this level) for STAIS, STAIT, UCLA-L or PSS
# - covid risk measure predicts less risk -taking
# model w/ covid risk (AIC = 73465.6) outperforms model with phase and day(AIC = 73516.2) and just phase (AIC =73516.4)


# TEMPORAL CONTEXT WITH COVID RISK:
if(!exists('model_pred_3timescales_covidRisk')){
  model_pred_3timescales_covidRisk = glmer(rdmChoice ~ 0 + rdmPOC1sc*rdmEarningSC + posShift +  rdmTrialSC + covq_PAB_q1_personalRisk_scaledNoNA + (1|subID), data=rdmGainQualtrics, family="binomial", offset=predmTask);
}

# AIC = 67154.3
# only covid risk not significant
# trial and earning opposite effects (both v sig)
# singular
#nObs = 97445

# TEMPORAL CONTEXT WITHOUT COVID RISK:
if(!exists('model_pred_3timescales')){
  model_pred_3timescales= glmer(rdmChoice ~ 0 + rdmPOC1sc*rdmEarningSC + posShift +  rdmTrialSC + (1|subID), data=rdmGainQualtrics, family="binomial", offset=predmTask);
}

# AIC = 67153.2 
# everything significant, positive effect of past outcome, earnings (positive) and trial (negative) have opposing effects, shift positive, interaction between past outcome and earnings negative
# singular
#nObs = 97445

```


Affective measures: 
# in set up, create subject level dataframe for affective measures (vs what we already have at the trial-level)
# How do we find what is unique across our qualtrics responses?
# could do a component type of analysis for all of the qualtrics responses (Gillan et al 2016 does something like this)
# we could do this in a variety of ways (across all responses v. across all SES responses)
# Gillan (2016) used a package in R to do it 
# "Try not to get paralyzed by the world of ways to do this stuff - start simple"
# Use QualtricsCombined_subID_scored to start - lets just get the summary and standard deviations 
# then we move toward correllogram / run factor analysis / PCA stuff
```{r}

```

```{r}
# save our models
save(file=outputfile, list=ls(pattern=glob2rx("model*"))); 
```

