---
title: "SES_COVQ_PCA"
author: "Hayley Brooks"
date: "12/1/2021"
output: html_document
---

This script will 1) create the long and wide individual-level dataframes (qualtrics exclusion applied) and 2) PCA analysis for SES and COVID questions


```{r setup, include=FALSE}
rm(list=ls())

library('config')
library('dplyr')
config = config::get()

# load qualtrics and exclusion 
exclsnPhs1_csv = file.path(config$path$combined, config$EXCLUSIONcsvs$RDM_AX_Qual_Phs1exclusion); # phase 1 exclusion path
exclsnPhs2_csv = file.path(config$path$combined, config$EXCLUSIONcsvs$RDM_AX_Qual_Phs2exclusion); # phase 2 exclusion path
qualtricsBothPhs_csv = file.path(config$path$combined, config$QUALTRICScsvs$Combined_subID_scored_noDuplicates); # qualtrics responses path

excludePhs1 = read.csv(exclsnPhs1_csv); # loads exclusion for phase 1
excludePhs2 = read.csv(exclsnPhs2_csv); # loads exclusion for phase 2
qualtricsBothPhs = read.csv(qualtricsBothPhs_csv); # load qualtrics responses

qualtricsBothPhs=arrange(qualtricsBothPhs, qualtricsBothPhs$subID); # subject IDs are not in order, rearranging dataframe so that subIDs are in order

# Remove the extra "X" column present as the first column in datasets
excludePhs1 = excludePhs1[,(2:ncol(excludePhs1))];
excludePhs2 = excludePhs2[,(2:ncol(excludePhs2))];
qualtricsBothPhs = qualtricsBothPhs[,(4:ncol(qualtricsBothPhs))]; # qualtrics has 3 columns of X variable
```

Apply Qualtrics Exclusions
```{r}

# Phase 1 Qualtrics: 3 participants excluded (but the 3rd, sub 461 is missing a response so here were are just putting NAs for two participants' qualtrics responses in phase 1)
subIDqualPhs1Exclude = excludePhs1$subID[!is.na(excludePhs1$qualPhs1exclude) & excludePhs1$qualPhs1exclude==1];

# Phase 2 Qualtrics: 1 participant excluded based on age response in phase
subIDqualPhs2Exclude = excludePhs2$subID[!is.na(excludePhs2$qualPhs2exclude) & excludePhs2$qualPhs2exclude==1];

# put NAs for responses (except for in the phase and subID columns)
qualtricsBothPhs[qualtricsBothPhs$subID %in% subIDqualPhs1Exclude & qualtricsBothPhs$phase==1,!colnames(qualtricsBothPhs) %in% c("phase","subID")] = NA; # phase1
qualtricsBothPhs[qualtricsBothPhs$subID %in% subIDqualPhs2Exclude & qualtricsBothPhs$phase==2,!colnames(qualtricsBothPhs) %in% c("phase","subID")] = NA; # phase2

# Note: if participants were excluded in phase 1, they are not automatically excluded in phase 2! Also, there are different exclusion for RDM, AX and Qualtrics.

```


```{r}
# Store sub IDs of those who are in phase 1 and phase 2 (post-exclusion)

# Participant IDs included in phase 1 RDM
Phs1subIDs = excludePhs1$subID[excludePhs1$qualPhs1exclude==0];
Phs1nSub = length(Phs1subIDs);

# Participant IDs included in phase 2 RDM
Phs2subIDs = excludePhs2$subID[!is.na(excludePhs2$qualPhs2exclude) & excludePhs2$qualPhs2exclude==0];
Phs2nSub = length(Phs2subIDs);

# Participant IDs included in both phases
BothPhsSubIDs = Phs2subIDs[Phs2subIDs %in% Phs1subIDs];
BothPhsnSub = length(BothPhsSubIDs)
```


There are a bunch of variables that we don't need to keep right now. Let's remove some of the columns and keep the variables that we need for PCA and correllograms:
```{r}
keepCols = c("subID", 
             "phase",
             "loc_fips",
             "stai_t_score",
             "stai_s_score",
             "pss_score",
             "pss_stressedToday",   
             "uclal_score",
             "covq_PAB_q1_personalRisk",
             "covq_PAB_q2_threat",
             "covq_PAB_q3_personallyDie",
             "covq_PAB_q4_otherPersonDie",
             "covq_PAB_q5_currentCases_recode",
             "covq_PAB_q6_tested_recode",
             "covq_PAB_q7_personalCovidSuspect_recode",
             "covq_PAB_q8_knowPositivePerson_recode",
             "covq_PAB_q9_socialDistanceLevel_recode",
             "demo_age",
             "demo_gender_recode",
             "demo_race_recode",
             "demo_ethnicity_recode",
             "ses_childhood_freeReducedLunch_recode",
             "ses_childhood_communityComp_recode",
             "ses_childhood_nationalComp_recode",
             "ses_motherEdLevel_recode",
             "ses_fatherEdLevel_recode",
             "ses_childhoood_homeOwnership_recode",
             "ses_current_billHelp_recode",
             "ses_current_mainResponsibilities_recode",
             "ses_personalEdLevel_recode",
             "ses_financialWorryFreq_recode",
             "ses_needbasedCollegeAid_recode");


# for the long individual-level dataframe, we want a row for each participant, even if they were excluded or did not participant (i.e. phase 2) because we want to keep these placeholders for the PCA so that rows and sub IDs line up
subLevelLong = as.data.frame(matrix(data=NA, nrow=nrow(excludePhs1)*2, ncol=length(keepCols), dimnames=list(c(NULL), c(keepCols))));
subLevelLong$subID=rep(1:544, each=2); # fill in sub ids
subLevelLong$phase=rep(1:2,times=2); # fill in phase

# distribute participants responses in subLevelLong
for (s in 1:nrow(subLevelLong)) {
  if (any(qualtricsBothPhs$subID==subLevelLong$subID[s] & qualtricsBothPhs$phase==subLevelLong$phase[s])){
      subLevelLong[s,3:length(keepCols)]= qualtricsBothPhs[qualtricsBothPhs$subID==subLevelLong$subID[s] & qualtricsBothPhs$phase==subLevelLong$phase[s],keepCols[3:length(keepCols)]]
  }
}



```


Create the subLevelWide dataframe (each participant gets a row)
```{r}
# use subLevelLong to make subLevelWide
keepColsWide =rep(keepCols, times=2)
subLevelWide = as.data.frame(matrix(data=NA, nrow = Phs1nSub, ncol = ncol(subLevelLong)*2, dimnames=list(c(NULL),c(keepColsWide)))); # column # and names will be adjusted when we delete redundancy between phases below.

for (s in 1:Phs1nSub) {
  sub = subLevelLong[which(subLevelLong$subID==Phs1subIDs[s]),]; # pull out one participant (using which because its doing funky stuff with NA otherwise)


  if(Phs1subIDs[s] %in% BothPhsSubIDs) { # if participant was in both phases
    subLevelWide[s,] = cbind(sub[1,],sub[sub$phase==2,][1,]); # in a single row, store phase 1 and phase 2 data for participant

  } else {
    subLevelWide[s,1:ncol(sub)] = sub[1,]; # in a single row, store phase 1 data for participant (the rest will be NaN)
  }

}

wideColNames = c("subID",
                 "phs1_phase",
                 "phs1_loc_fips",
                 "phs1_stai_t_score",
                 "phs1_stai_s_score",
                 "phs1_pss_score",
                 "phs1_pss_stressedToday", 
                 "phs1_uclal_score",
                 "phs1_covq_PAB_q1_personalRisk",
                 "phs1_covq_PAB_q2_threat",
                 "phs1_covq_PAB_q3_personallyDie",
                 "phs1_covq_PAB_q4_otherPersonDie",
                 "phs1_covq_PAB_q5_currentCases_recode",
                 "phs1_covq_PAB_q6_tested_recode",
                 "phs1_covq_PAB_q7_personalCovidSuspect_recode",
                 "phs1_covq_PAB_q8_knowPositivePerson_recode",
                 "phs1_covq_PAB_q9_socialDistanceLevel_recode",
                 "demo_age",
                 "demo_gender_recode",
                 "demo_race_recode",
                 "demo_ethnicity_recode",
                 "phs1_ses_childhood_freeReducedLunch_recode",
                 "phs1_ses_childhood_communityComp_recode",
                 "phs1_ses_childhood_nationalComp_recode",
                 "phs1_ses_motherEdLevel_recode",
                 "phs1_ses_fatherEdLevel_recode",
                 "phs1_ses_childhoood_homeOwnership_recode",
                 "phs1_ses_current_billHelp_recode",
                 "phs1_ses_current_mainResponsibilities_recode",
                 "phs1_ses_personalEdLevel_recode",
                 "phs1_ses_financialWorryFreq_recode",
                 "phs1_ses_needbasedCollegeAid_recode",
                 "phs2_subID",
                 "phs2_phase",
                 "phs2_loc_fips",
                 "phs2_stai_t_score",
                 "phs2_stai_s_score",
                 "phs2_pss_score",
                 "phs2_pss_stressedToday", 
                 "phs2_uclal_score",
                 "phs2_covq_PAB_q1_personalRisk",
                 "phs2_covq_PAB_q2_threat",
                 "phs2_covq_PAB_q3_personallyDie",
                 "phs2_covq_PAB_q4_otherPersonDie",
                 "phs2_covq_PAB_q5_currentCases_recode",
                 "phs2_covq_PAB_q6_tested_recode",
                 "phs2_covq_PAB_q7_personalCovidSuspect_recode",
                 "phs2_covq_PAB_q8_knowPositivePerson_recode",
                 "phs2_covq_PAB_q9_socialDistanceLevel_recode",
                 "phs2_demo_age",
                 "phs2_demo_gender_recode",
                 "phs2_demo_race_recode",
                 "phs2_demo_ethnicity_recode",
                 "phs2_ses_childhood_freeReducedLunch_recode",
                 "phs2_ses_childhood_communityComp_recode",
                 "phs2_ses_childhood_nationalComp_recode",
                 "phs2_ses_motherEdLevel_recode",
                 "phs2_ses_fatherEdLevel_recode",
                 "phs2_ses_childhoood_homeOwnership_recode",
                 "phs2_ses_current_billHelp_recode",
                 "phs2_ses_current_mainResponsibilities_recode",
                 "phs2_ses_personalEdLevel_recode",
                 "phs2_ses_financialWorryFreq_recode",
                 "phs2_ses_needbasedCollegeAid_recode"); #This includes redundancies that will be removed!

colnames(subLevelWide) = wideColNames; # add column names

discardCols = c("phs1_phase",
                "phs2_subID",
                "phs2_phase",
                "phs2_demo_race_recode",
                "phs2_demo_ethnicity_recode",
                "phs2_demo_gender_recode",
                "phs2_demo_age"); # columns that will be discarded for redundancy

subLevelWide = subLevelWide[!names(subLevelWide) %in% discardCols]; # discard columns
```




Scale affective variables by max values for each affective measure
```{r}
# long dataframe
# pss, stait, uclal
subLevelLong$pss_score_scaled = subLevelLong$pss_score/max(subLevelLong$pss_score, na.rm = T);
subLevelLong$stai_t_score_scaled = subLevelLong$stai_t_score/max(subLevelLong$stai_t_score, na.rm = T);
subLevelLong$stai_s_score_scaled = subLevelLong$stai_s_score/max(subLevelLong$stai_s_score, na.rm = T);
subLevelLong$uclal_score_scaled = subLevelLong$uclal_score/max(subLevelLong$uclal_score, na.rm = T);

# covid risk
subLevelLong$covq_PAB_q1_personalRisk_scaled = (subLevelLong$covq_PAB_q1_personalRisk-1)/max(subLevelLong$covq_PAB_q1_personalRisk-1, na.rm = T)

subLevelLong$covq_PAB_q1_personalRisk_scaledNoNA = subLevelLong$covq_PAB_q1_personalRisk_scaled;
subLevelLong$covq_PAB_q1_personalRisk_scaledNoNA[is.na(subLevelLong$covq_PAB_q1_personalRisk_scaledNoNA)] = 0;

# wide dataframe
# pss, stait, uclal
# phase 1
subLevelWide$phs1_pss_score_scaled = subLevelWide$phs1_pss_score/max(subLevelWide$phs1_pss_score, na.rm = T);
subLevelWide$phs1_stai_t_score_scaled = subLevelWide$phs1_stai_t_score/max(subLevelWide$phs1_stai_t_score, na.rm = T);
subLevelWide$phs1_stai_s_score_scaled = subLevelWide$phs1_stai_s_score/max(subLevelWide$phs1_stai_s_score, na.rm = T);
subLevelWide$phs1_uclal_score_scaled = subLevelWide$phs1_uclal_score/max(subLevelWide$phs1_uclal_score, na.rm = T);

# phase 2
subLevelWide$phs2_pss_score_scaled = subLevelWide$phs2_pss_score/max(subLevelWide$phs2_pss_score, na.rm = T);
subLevelWide$phs2_stai_t_score_scaled = subLevelWide$phs2_stai_t_score/max(subLevelWide$phs2_stai_t_score, na.rm = T);
subLevelWide$phs2_stai_s_score_scaled = subLevelWide$phs2_stai_s_score/max(subLevelWide$phs2_stai_s_score, na.rm = T);
subLevelWide$phs2_uclal_score_scaled = subLevelWide$phs2_uclal_score/max(subLevelWide$phs2_uclal_score, na.rm = T);


# covid risk
# phase 1
subLevelWide$phs1_covq_PAB_q1_personalRisk_scaled = (subLevelWide$phs1_covq_PAB_q1_personalRisk-1)/max(subLevelWide$phs1_covq_PAB_q1_personalRisk-1, na.rm = T);

# replace NA with 0
subLevelWide$phs1_covq_PAB_q1_personalRisk_scaledNoNA = subLevelWide$phs1_covq_PAB_q1_personalRisk_scaled;
subLevelWide$phs1_covq_PAB_q1_personalRisk_scaledNoNA[is.na(subLevelWide$phs1_covq_PAB_q1_personalRisk_scaledNoNA)] = 0;

# phase 2
subLevelWide$phs2_covq_PAB_q1_personalRisk_scaled = (subLevelWide$phs2_covq_PAB_q1_personalRisk-1)/max(subLevelWide$phs2_covq_PAB_q1_personalRisk-1, na.rm = T);

# replace NA with 0
subLevelWide$phs2_covq_PAB_q1_personalRisk_scaledNoNA = subLevelWide$phs2_covq_PAB_q1_personalRisk_scaled;
subLevelWide$phs2_covq_PAB_q1_personalRisk_scaledNoNA[is.na(subLevelWide$phs2_covq_PAB_q1_personalRisk_scaledNoNA)] = 0;

```



#### Can we find what is unique across our qualtrics responses for SES and Covid Q?
# Correlograms
```{r}
# Look at relationship between affective measures (STAIT, STAIS, PSS, UCLAL <--> perceived COVID risk) (overall correlogram)

library("GGally"); # load package we need for correlogram

# phase one only - all affective measures
tmpPhs1 = subLevelWide[names(subLevelWide) %in% c("phs1_stai_s_score_scaled", "phs1_stai_t_score_scaled", "phs1_pss_score_scaled", "phs1_pss_stressedToday", "phs1_uclal_score_scaled", "phs1_covq_PAB_q1_personalRisk_scaled")];
ggpairs(tmpPhs1, title = "Phase 1 correlogram", columnLabels = c("state anxiety", "trait anxiety", "stress", "stress today", "loneliness", "covid risk"));

# phase two only - all affective measures
tmpPhs2 = subLevelWide[names(subLevelWide) %in% c("phs2_stai_s_score_scaled", "phs2_stai_t_score_scaled", "phs2_pss_score_scaled", "phs2_pss_stressedToday", "phs2_uclal_score_scaled", "phs2_covq_PAB_q1_personalRisk_scaled")];
ggpairs(tmpPhs2, title = "Phase 2 correlogram", columnLabels = c("state anxiety", "trait anxiety", "stress", "stress today", "loneliness", "covid risk"));

# phase 1 and 2 covid risk
tmpCovidRisk = subLevelWide[names(subLevelWide) %in% c("phs1_covq_PAB_q1_personalRisk_scaled", "phs2_covq_PAB_q1_personalRisk_scaled")];
ggpairs(tmpCovidRisk, title = "covid risk between phases", columnLabels = c("phs 1 covid risk", "phs 2 covid risk"));

# difference in covid risk across phases (just plot it)
subLevelWide$covidRiskDiff = subLevelWide$phs1_covq_PAB_q1_personalRisk_scaled - subLevelWide$phs2_covq_PAB_q1_personalRisk_scaled;
hist(subLevelWide$covidRiskDiff, main=sprintf("covid risk difference\n mean = %.5s, range = %s - %s", mean(subLevelWide$covidRiskDiff, na.rm=T), min(subLevelWide$covidRiskDiff, na.rm = T), c(max(subLevelWide$covidRiskDiff,na.rm=T)) ), xlab = "phase 1 - phase 2", ylab="number of people", breaks=10);
# 113 people have covidriskDiff = 0 (histogram shows the 113 below zero)

# how does change in covid risk relate to phase 2 (end-state affect)
tmpPhs2covidDiff = subLevelWide[names(subLevelWide) %in% c("phs2_stai_s_score_scaled", "phs2_stai_t_score_scaled", "phs2_pss_score_scaled", "phs2_pss_stressedToday", "phs2_uclal_score_scaled", "covidRiskDiff")];
ggpairs(tmpPhs2covidDiff, title = "Phase 2 correlogram with covid risk diff", columnLabels = c("state anxiety", "trait anxiety", "stress", "stress today", "loneliness", "covid risk diff"));
# not seeing a relationship between change covid risk and end state affect (phase 2)


# difference in stait across phases (just plot it)
staitDiff = subLevelWide$phs1_stai_t_score_scaled - subLevelWide$phs2_stai_t_score_scaled
hist(staitDiff, main=sprintf("stait difference\n mean = %.5s, range = %s - %s", mean(staitDiff, na.rm=T), min(staitDiff, na.rm = T), c(max(staitDiff,na.rm=T)) ), xlab = "phase 1 - phase 2", ylab="number of people", breaks=10);

# difference in stais across phases (just plot it)
staisDiff = subLevelWide$phs1_stai_s_score_scaled - subLevelWide$phs2_stai_s_score_scaled
hist(staisDiff, main=sprintf("stais difference\n mean = %.5s, range = %s - %s", mean(staisDiff, na.rm=T), min(staisDiff, na.rm = T), c(max(staisDiff,na.rm=T)) ), xlab = "phase 1 - phase 2", ylab="number of people", breaks=10);

# difference in pss across phases (just plot it)
pssDiff = subLevelWide$phs1_pss_score_scaled - subLevelWide$phs2_pss_score_scaled
hist(pssDiff, main=sprintf("pss difference\n mean = %.5s, range = %s - %s", mean(pssDiff, na.rm=T), min(pssDiff, na.rm = T), c(max(pssDiff,na.rm=T)) ), xlab = "phase 1 - phase 2", ylab="number of people", breaks=10);

# difference in uclal across phases (just plot it)
uclalDiff = subLevelWide$phs1_uclal_score_scaled - subLevelWide$phs2_uclal_score_scaled
hist(uclalDiff, main=sprintf("uclal difference\n mean = %.5s, range = %s - %s", mean(uclalDiff, na.rm=T), min(uclalDiff, na.rm = T), c(max(uclalDiff,na.rm=T)) ), xlab = "phase 1 - phase 2", ylab="number of people", breaks=10);
```


# SES questions (there are 11 of them in both phases)
```{r}
# how does change in covid risk relate to phase 2 (end-state affect)
sesDFphs1= subLevelWide[names(subLevelWide) %in% c("phs1_ses_childhood_freeReducedLunch_recode",
                                               "phs1_ses_childhood_communityComp_recode",
                                               "phs1_ses_childhood_nationalComp_recode",
                                               "phs1_ses_motherEdLevel_recode",
                                               "phs1_ses_fatherEdLevel_recode" ,
                                               "phs1_ses_childhoood_homeOwnership_recode" ,
                                               "phs1_ses_personalEdLevel_recode",
                                               "phs1_ses_financialWorryFreq_recode",
                                               "phs1_ses_needbasedCollegeAid_recode")];

ggpairs(sesDFphs1, title = "Phase 1 correlogram with SES", columnLabels = c("reduced lunch", "communityComp", "nationalComp", "momED", "dadED", "childHomeOwn", "personalEDlev", "worryFreq", "collegeAid"));


sesDFphs2= subLevelWide[names(subLevelWide) %in% c("phs2_ses_childhood_freeReducedLunch_recode",
                                               "phs2_ses_childhood_communityComp_recode",
                                               "phs2_ses_childhood_nationalComp_recode",
                                               "phs2_ses_motherEdLevel_recode",
                                               "phs2_ses_fatherEdLevel_recode" ,
                                               "phs2_ses_childhoood_homeOwnership_recode" ,
                                               "phs2_ses_personalEdLevel_recode",
                                               "phs2_ses_financialWorryFreq_recode",
                                               "phs2_ses_needbasedCollegeAid_recode")];

ggpairs(sesDFphs2, title = "Phase 2 correlogram with SES", columnLabels = c("reduced lunch", "communityComp", "nationalComp", "momED", "dadED", "childHomeOwn", "personalEDlev", "worryFreq", "collegeAid"));


# Removed 1) current bill help because it was very unrelated to the others which could be the result of differences in the way participants interpreted the question and 2) current responsibilities because it categorical or ordered
# Phase 1 and phase 2 broadly similar

#looking at differences in responses across phases:
# need to plot this so we can see how many points were different:
for (c in 1:ncol(sesDFphs1)) {
  plot(sesDFphs1[,c],sesDFphs2[,c])
}


# Next: moving toward "sophisticated", data-driven way would be to move toward PCA with a "smartly-weighted" combination of the components.
# is covid risk related to any of the affective measures? - answered above
# is change in covid risk related to change in affective measures - not done yet

```


PCA!
```{r}
# PCA for SES (using 9 out of 11 SES measures - see above for why we removed two)

sesPhs1pca = prcomp(na.omit(sesDFphs1), center=TRUE, scale = TRUE);
#Importance of components:
#                          PC1    PC2    PC3     PC4     PC5     PC6     PC7     PC8    PC9
#Standard deviation     1.7582 1.0707 1.0016 0.91772 0.89793 0.81304 0.77360 0.70472 0.5955
#Proportion of Variance 0.3435 0.1274 0.1115 0.09358 0.08959 0.07345 0.06649 0.05518 0.0394
#Cumulative Proportion  0.3435 0.4709 0.5823 0.67589 0.76548 0.83893 0.90542 0.96060 1.0000


sesPhs2pca = prcomp(na.omit(sesDFphs2), center=TRUE, scale = TRUE);

# Importance of components:
#                           PC1    PC2    PC3     PC4     PC5     PC6     PC7     PC8     PC9
# Standard deviation     1.7882 1.0755 1.0086 0.91718 0.83718 0.82613 0.74987 0.68966 0.60498
# Proportion of Variance 0.3553 0.1285 0.1130 0.09347 0.07787 0.07583 0.06248 0.05285 0.04067
# Cumulative Proportion  0.3553 0.4838 0.5968 0.69030 0.76817 0.84401 0.90649 0.95933 1.00000

# the components in both phases are very similar
# for example, comparing PC1 in phase 1 and 2:
plot(as.vector(sesPhs1pca$rotation[,1]),as.vector(sesPhs2pca$rotation[,1]));



# ADD SES PCA components to individual-level dataframes
# This will take a little thinking because there are 520 PCAs (participants who did not respond to some SES questions are skipped)
# Then it will be a little tricky adding phase 1 and phase 2 components to the long dataframe since participants have 1-2 rows.

# accessing component for each person looks like:
sesCompPhs1 = sesPhs1pca$x[,'PC1'] # 520 rows
sesCompPhs1Names = as.numeric(names(sesCompPhs1)); # store the names for each value (i.e. participant ID)
sesCompPhs2 = sesPhs2pca$x[,'PC1'] # 347 rows
sesCompPhs2Names = as.numeric(names(sesCompPhs2)); # store the names for each value (i.e. participant ID)

# for phase 1, there are PC1 for 517 participants (3 excluded and 21 who did not respond to all SES questions)
for (s in 1:length(sesCompPhs1Names)){
  subLevelWide$phs1_sesPCA[subLevelWide$subID ==sesCompPhs1Names[s]] = sesCompPhs1[s]; # store the PC1 value 
  subLevelLong$sesPCA[subLevelLong$subID == sesCompPhs1Names[s] & subLevelLong$phase==1] = sesCompPhs1[s]; # store the PC1 value 
}

# for phase 2, there are PC1 for 347 participants (1 participant excluded and 8 who did not respond to all SES questions)
for (s in 1:length(sesCompPhs2Names)){
  subLevelWide$phs2_sesPCA[subLevelWide$subID ==sesCompPhs2Names[s]] = sesCompPhs2[s]; # store the PC2 value 
  subLevelLong$sesPCA[subLevelLong$subID == sesCompPhs2Names[s] & subLevelLong$phase==2] = sesCompPhs2[s]; # store the PC2 value
}



```


Look at corellograms and do PCA with covid questions
```{r}
# COVID-19 QUESTIONS questions (there are 9 of them in both phases)
# phase 1:
covqDFphs1= subLevelWide[names(subLevelWide) %in% c("phs1_covq_PAB_q1_personalRisk_scaled",
                                               "phs1_covq_PAB_q2_threat",
                                               "phs1_covq_PAB_q3_personallyDie",
                                               "phs1_covq_PAB_q4_otherPersonDie",
                                               "phs1_covq_PAB_q5_currentCases_recode" ,
                                               "phs1_covq_PAB_q6_tested_recode" ,
                                               "phs1_covq_PAB_q7_personalCovidSuspect_recode",
                                               "phs1_covq_PAB_q8_knowPositivePerson_recode",
                                               "phs1_covq_PAB_q9_socialDistanceLevel_recode")];

ggpairs(covqDFphs1, title = "Phase 1 correlogram COVID Qs", columnLabels = c("threat", "personal death", "death of other", "current cases", "tested", "suspected covid", "know positive person", "social distance", "personal risk"));


# phase 2:
covqDFphs2= subLevelWide[names(subLevelWide) %in% c("phs2_covq_PAB_q1_personalRisk_scaled",
                                               "phs2_covq_PAB_q2_threat",
                                               "phs2_covq_PAB_q3_personallyDie",
                                               "phs2_covq_PAB_q4_otherPersonDie",
                                               "phs2_covq_PAB_q5_currentCases_recode" ,
                                               "phs2_covq_PAB_q6_tested_recode" ,
                                               "phs2_covq_PAB_q7_personalCovidSuspect_recode",
                                               "phs2_covq_PAB_q8_knowPositivePerson_recode",
                                               "phs2_covq_PAB_q9_socialDistanceLevel_recode")];

ggpairs(covqDFphs2, title = "Phase 2 correlogram COVID Qs", columnLabels = c("threat", "personal death", "death of other", "current cases", "tested", "suspected covid", "know positive person", "social distance", "personal risk"));

covqDiff = covqDFphs2-covqDFphs1; # take the difference in covid qs between phase 2 and phase 1

covqDifftopFive = covqDiff[names(covqDiff) %in% c("phs2_covq_PAB_q1_personalRisk_scaled",
                                               "phs2_covq_PAB_q2_threat",
                                               "phs2_covq_PAB_q3_personallyDie",
                                               "phs2_covq_PAB_q4_otherPersonDie",
                                           "phs2_covq_PAB_q8_knowPositivePerson_recode")];

colnames(covqDifftopFive) = c("threatDiff", "personallyDieDiff", "otherPersonDieDiff", "knowPosPersonDiff","personalRiskDiff");

ggpairs(covqDifftopFive, title = "Diff correlogram COVID Qs", columnLabels = c("threat", "personal death", "death of other","know positive person","personal risk"));



# The top five covid qs that stand out are personal risk, threat, personal death, death of other and know any one who tested positive. However, the know anyone who tested positive is coded funky and its unclear whats happening there. The change across phases for the top four (throwing out know anyone who tested positive right now) is pretty small, but the changes in personal risk distribution are related to RDM and the distribution of changes across phases is widest (ie it has the biggest change) and may be the variable to go with. The top four variables are highly related and the degree to which they change across phases is also correlated. 


mean(covqDifftopFive$personalRiskDiff, na.rm=T); #-0.03319209*6 = -0.1991525 (to match the scale with the following cov qs becase personal risk is on a 0-1 scale)
mean(covqDifftopFive$threatDiff, na.rm=T); # -0.07932011
mean(covqDifftopFive$personallyDieDiff, na.rm=T); # 0.04519774
mean(covqDifftopFive$otherPersonDieDiff, na.rm=T); # -0.115493
mean(covqDifftopFive$knowPosPersonDiff, na.rm=T); # 0.03380282
# largest difference is with personal risk, followed by other personally die

var(covqDifftopFive$personalRiskDiff*6, na.rm=T); #1.687563 (multiplied score by 6 because math for variance is more complicated)
var(covqDifftopFive$threatDiff, na.rm=T); # 1.249372
var(covqDifftopFive$personallyDieDiff, na.rm=T); # 1.176422
var(covqDifftopFive$otherPersonDieDiff, na.rm=T); #  1.899053
var(covqDifftopFive$knowPosPersonDiff, na.rm=T); # 0.7559163

# highest variance is other person die followed by personal risk ( similar)
# moving toward rationalizing why we should use to personal risk (While also accounting for other die?)
# Change in personal risk and death of other are not as correlated as we'd think given their relationship within each phase. This could be that the change is just noise and so there is no meaningful shift that is happening. 


```

```{r}
# PCA for COVID Q (using 4 out of 9 covid q measures - see above for why we are going with these)
# personal risk, threat, personal death, death of other

covqDFphs1top4 = covqDFphs1[names(covqDFphs1) %in% c("phs1_covq_PAB_q1_personalRisk_scaled",
                                               "phs1_covq_PAB_q2_threat",
                                               "phs1_covq_PAB_q3_personallyDie",
                                               "phs1_covq_PAB_q4_otherPersonDie")];

covqDFphs2top4 = covqDFphs2[names(covqDFphs2) %in% c("phs2_covq_PAB_q1_personalRisk_scaled",
                                               "phs2_covq_PAB_q2_threat",
                                               "phs2_covq_PAB_q3_personallyDie",
                                               "phs2_covq_PAB_q4_otherPersonDie")];

covqPhs1pca = prcomp(na.omit(covqDFphs1top4), center=TRUE, scale = TRUE);

#                                             PC1        PC2         PC3        PC4
# phs1_covq_PAB_q2_threat              -0.4836368  0.7337859 -0.02583329 -0.4764307
# phs1_covq_PAB_q3_personallyDie       -0.4917736 -0.6281408  0.35491237 -0.4874783
# phs1_covq_PAB_q4_otherPersonDie      -0.4851348 -0.2288239 -0.82351128  0.1846970
# phs1_covq_PAB_q1_personalRisk_scaled -0.5374927  0.1209832  0.44181334  0.7080012

# Importance of components:
#                           PC1    PC2    PC3     PC4
# Standard deviation     1.5437 0.8138 0.7599 0.61433
# Proportion of Variance 0.5957 0.1656 0.1444 0.09435
# Cumulative Proportion  0.5957 0.7613 0.9056 1.00000



covqPhs2pca = prcomp(na.omit(covqDFphs2top4), center=TRUE, scale = TRUE);

#                                             PC1        PC2          PC3        PC4
# phs2_covq_PAB_q2_threat              -0.4498989  0.8084514 -0.001388992 -0.3794673
# phs2_covq_PAB_q3_personallyDie       -0.5038844 -0.5600893 -0.280316427 -0.5948304
# phs2_covq_PAB_q4_otherPersonDie      -0.5006275 -0.1801365  0.821079910  0.2067626
# phs2_covq_PAB_q1_personalRisk_scaled -0.5413536  0.0160342 -0.497240941  0.6778131

# Importance of components:
#                           PC1    PC2    PC3     PC4
# Standard deviation     1.5423 0.8440 0.7358 0.60619
# Proportion of Variance 0.5947 0.1781 0.1353 0.09187
# Cumulative Proportion  0.5947 0.7728 0.9081 1.00000

# similar pattern in phase 2


# Component one weighs each of the four components roughly equally and accounts for 60% of the variance. This could mean that picking one variable (i.e. personal risk) would be ok and the argument for would be that its more straightforward and more interpretable. The counter argument would be to use the weighted average of the four variables.

# Going ahead using personal risk variable for now.


```

Save individual-level data frames
```{r}

subLongFilePath = file.path(config$path$Rdata, config$Rdata_files$QualtricsSubLevelLong);
subWideFilePath = file.path(config$path$Rdata, config$Rdata_files$QualtricsSubLevelWide);
save(file= subLongFilePath, subLevelLong);
save(file= subWideFilePath, subLevelWide);
```

